GWAS datasets for FDR benchmarking

### 2017-09-13 datasets/GWAS/GWAS.Rmd
remove chrono study; add asthma study

### GWAS for Self-reported Chronotype

A [study by Hu et al. (Nature Comm, 2016)](https://www.ncbi.nlm.nih.gov/pubmed/?term=26835600) presented a GWAS of 89,283 individuals that identified genetic variants associated with self-reporting of being a morning person.

Let's download the data from the `ebi.ac.uk` ftp site:

```{bash, GWAS2-download, eval=FALSE}
ftp ftp://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/
cd KhorCC_27064256_GCST003467
get "Khor-27064256.txt"
exit
```
Next, we'll read the `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS2-verify-contents}
glauc <- fread("Khor-27064256.txt", header=TRUE)
dim(glauc)
head(glauc)
```

@@

- `pvalue`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
  - `gt.rate`: ?
  - `freq.b`: ?
  - `dose.b`: ?

for 10,000 SNPs - unfortunately, looks like they only give us the *top* 
10K SNPs, so this is not useful for our purposes.

```{r, GWAS2-cleanup}
rm(chrono)
```

Let's download the data from the Gabriel site

```{bash, GWAS2-download, eval=FALSE}
curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
```

>>>>

[This study by Moffatt et al. (2010, NEJM)](http://www.nejm.org/doi/full/10.1056/NEJMoa0906312?query=TOC) performed a large
GWAS on  10,365 people with asthma and 16,110 controls 23 individual studies.

Let's download the data from the Gabriel site. The second file contains metadata
for the columns in the main data file.

```{bash, GWAS2-download, eval=FALSE}
curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
curl -O "https://beaune.cng.fr/gabriel/gabriel_results_description.xls"
```

Next we unzip the file and remove the zipped copy.

```{bash, GWAS2-unzip, eval=FALSE}
unzip gabriel_results.zip
rm gabriel_results.zip
```

Next, we'll read in the unzipped `.txt` file and the results desription `.xls` 
file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS2-verify-contents}
# read in meta-data
library(readxl)
meta <- read_xls("gabriel_results_description.xls")
head(meta)
dim(meta)

# read in data
asthma <- fread("gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt", header=TRUE)
dim(asthma)

# verify meta-data rows = data cols
nrow(meta) == ncol(asthma)

# remove study-specific columns
library(dplyr)
empty <- which(!is.na(meta$description))
asthma <- asthma %>%
            select(empty)
meta <- meta %>% 
            slice(empty)
head(asthma)
```

@@

- `P_fix` and `P_ran`:p-values for fixed and random effect models
- `theta_fix` and `theta_fix`:effect size for fixed and random effect models
- `se_theta_fix` and `se_theta_fix`:standard error for fixed and random effect models
- additional covariates: 
    - `no_studies_meta`: Number of studies included in the meta-analysis with this SNP
    - `freq_all_1_max` and `freq_all_1_min`: Maximum and Minimum Allele Frequency of Allele 1 across samples

for 567,589 SNPs.

```{r, GWAS2-cleanup}
rm(asthma)
```
